---
title: "Exploratory Data Analysis"
author: "Achmad Wildan Al Aziz"
date: "October 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DT)
library(ggplot2)
library(corrplot)
library(dplyr)
```

Exploratory data analysis(EDA) is an approach to data analysis for summarising and visualising the important characteristics of a data set. EDA is not a formal process with a strict set of rules. EDA is an important part of any data analysis because you always need to investigate the quality of your data. 

#### Objective
Your goal during EDA is to develop an understanding of your data. 

- Discover any existing patterns
- Identify mistakes
- Suggest hypothesis that can be tested
- Check assumptions
- Identify relationships among explanatory variables
- Assess direction and size of relationship between expplanatory and outcome variables

***
### Example of EDA

#### Univariate

##### Univariate Quantitative

###### Measures of central tendancy

- Mean
- Median
- Mode

###### Measures of dispersion

- Min
- Max
- Range
- Quartiles
- Variance
- Standard deviation

###### Other measures include

- Skewness
- Kurtosis

##### Univariate Graphical

- Histogram
- Box plots
- Bar plots
- Kernel density plots
- Bivariate
- Bivariate Quantitative
- Bivariate analysis include:

- Crosstabs
- Covariance
- Correlation

- Cluster analysis
- Analysis of variance (ANOVA)
- Factor analysis
- Principal component analysis (PCA)
- Bivariate Graphical
- Graphical techniques include:

- Scatterplot
- Box plot

### Diamonds Dataset

Merupakan data harga berlian dan karakteristiknya.
Variabel yang digunakan adalah sebagai berikut

Variabel | Keterangan
--------- | ------------------------------------------------------------------------------------------
price | _price in_ ($US) ($326-$18,823)
carat | _weight of the diamond_ (0.2-5.01)
cut | _quality of the cut_ (Fair, Good, Very Good, Premium, Ideal)
color | _diamond colour_,  from J (worst) to D (best)
clarity | _a measurement of how clear the diamond is_ (I1 (worst), SI2, SI1, VS2, VS1, VVS2, VVS1, IF (best))
x | _length in mm_ (0-10.74)
y | _width in mm_ (0-58.9)
z | _depth in mm_ (0-31.8)
depth | _total depth percentage_ = z / mean(x, y) = 2 * z / (x + y), (43-79)
table | _width of top of diamond relative to widest point_ (43-95)

```{r Data, message=FALSE, warning=FALSE}
data("diamonds")
datatable(diamonds)
```

```{r}
summary(diamonds)
```

##### Missing values

grafik berikut untuk mengecej _missing values_
```{r}
library(naniar)
gg_miss_var(diamonds, show_pct = TRUE)
```

tidak ada missing values pada data _diamonds_

##### Visualization {.tabset}

Berikut adalah korelasi antar variabel dari data _diamonds_ 
```{r}
num=diamonds[c('price','carat','x','y','z','depth','table')]
corrplot(cor(num),type="full",method="square")
```

beberapa variabel terjadi multikolinearitas, yaitu variabel carat, x, y, z. Variabel _price_ berkorelasi positif cukup tinggi dengan variabel carat, x, y, z.

###### Price

berikut adalah histogram untuk variabel respon, _price_.
```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price, fill=..count..), binwidth = 900)+
    scale_y_continuous(name = "Frequency") +
    scale_x_continuous(name = " Price ($US)") +
    ggtitle("Frequency histogram Price of diamonds ($US)")
```

Tampak histogram _skew kanan_, sehingga variabel _price_ tidak berdistribusi normal. Mayoritas produsen memberikan harga sekitar 20000-4000 ($US). Sebagian besar Berlian harganya kurang dari $US 5000. Varibel Price dapat dilakukan transformasi.

###### carat

berikut adalah histogram untuk variabel, _carat_.
```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat, fill=..count..), binwidth = 0.4)+
    scale_y_continuous(name = "carat") +
    scale_x_continuous(name = " carat of the diamond") +
    ggtitle("Frequency histogram carat of the diamond")
```

Tampak histogram _skew kanan_, sehingga variabel _carat_ tidak berdistribusi normal. Variabel _carat_ bernilai antara 0,2 hingga 5.01, tampak pada histogram _carat_ yang nilainya lebih dari 3 tidak muncul karena frekuensinya terlalu sedikit. Dapat dilakukan penyesuaian seperti histogram berikut.

```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat, fill=..count..), binwidth = 0.4)+
    scale_y_continuous(name = "Frequency") +
    scale_x_continuous(name = " weight of the diamond") +
    ggtitle("Frequency histogram weight of the diamond") +
    coord_cartesian(ylim = c(0, 100))
```

Atau dapat juga ditabelkan seperti tabel berikut.

```{r message=FALSE, warning=FALSE}
diamonds %>% count(cut_width(carat, 0.4))
```

###### Boxplot

Berikut adalah Boxplot untuk setiap variabel numerik

```{r message=FALSE, warning=FALSE}
require(reshape2)
ggplot(data = melt(num[,-1]), aes(x=variable, y=value)) +
  geom_boxplot() +
  facet_wrap(~variable, scales='free')
```

juga dapat ditampilkan Boxplot untuk variabel kategorik

```{r message=FALSE, warning=FALSE}
kat=diamonds[c('cut','color','clarity','price')]

ggplot(data = kat, aes(x=cut, y=price)) +
  geom_boxplot() 
```
